---
title: "malletparseDSI"
author: "Arthur Koehl"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
---

```{r setup, include = FALSE}
  knitr::opts_chunk$set(
			collapse = TRUE,
			comment = "#>"
  )
```

# Quick Start

This page outlines the ways to use this package for parsing various output files from Mallet.  
Refer the the README for this package for more information.
Note that topic ids and document ids start at 0 in Mallet's outputfiles.

first load the package
```{r}
library(malletparseDSI)
```

## 1. State File

The State file contains all the information that you will need from the model.  
This command shows how to parse the topic state file without saving the doc topics to an output file. 
Note that with the 'dtflag' argument set to 0, the document topics matrix does not get saved in R and will be empty in the returned object.
Set 'dtflag' to 1 to save the matrix, this argument is independent of saving the doc topics to file (i.e can be set to 1 or 0 when you save to file)

```{r}
# use this to save to file: statelist = rcpp_parse_topic_state("../data/state-file.gz", dtflag=0, dtfile="../data/doc-topics.gz") 
statefile = system.file("extdata", "state-file.gz", package="malletparseDSI")
statelist = rcpp_parse_topic_state(statefile, dtflag=0)
str(statelist)
```

## 2. Doc Topics File

In Mallet, you can output the document topics matrix to a file. 
When parsing the file with **rcpp_parse_doc_topics** you can pass a vector of integers of the length of each document in the model. 
If you pass that vector, rcpp_parse_doc_topics will calculate the topic frequencies, a necessary part of **LDAvis**.

```{r}
dtfile = system.file("extdata", "doc-topics.dat", package="malletparseDSI")
dtlist = rcpp_parse_doc_topics(dtfile, 20, doc_lens = statelist$doc_lens) #second arg is number of documents to save for each topic
str(dtlist)
```

You can do exactly the same thing with a gzipped doc topics file created with the **rcpp_parse_topic_state** function.

```{r}
dtfile2 = system.file("extdata", "doc-topics.gz", package="malletparseDSI")
dtlist2 = rcpp_parse_doc_topics(dtfile2, 20, doc_lens = statelist$doc_lens)
str(dtlist2)
```

## 3. Topic Terms

### a. Word Topic Counts File 

With this package you can extract the topic terms matrix wihtout using the Mallet state file by using the information from word-topic-counts or word-weights.  


To extract the topic terms from the word-topic-counts file, you need to pass in the number of topics as an input parameter.
```{r}
wtcfile = system.file("extdata", "word-topic-counts.dat", package="malletparseDSI")
topic_terms = rcpp_parse_word_topic_counts(wtcfile, 10) # third arg is ntopics
print(dim(topic_terms))
print(head(topic_terms[,1:5]))
```

## b. Topic Word Weights File 

Just like above, but you don't need to pass the number of topics as an input parameter.
```{r}
wwfile = system.file("extdata", "word-weights.dat", package="malletparseDSI")
topic_terms = rcpp_parse_topic_word_weights(wwfile)
print(dim(topic_terms))
print(head(topic_terms[,1:5]))
```

## LDAvis 

This package has a modified version of LDAvis' createJSON function that does not require the document topics matrix as input.
Instead, it uses the topic_frequencies calculated from the state file, or from a sample of the document topics.
This avoids having to ever load the document topics matrix into memory.

### a. modified createJSON
```{r}
tt_dist = t(apply(topic_terms, 1, function(x) x/sum(x))) ## convert topic terms count matrix to distribution

json = no_doc_topics_createJSON(
	phi = tt_dist, 
	topic.frequency = dtlist$topic_frequencies, 
	vocab = statelist$terms, 
	term.frequency = statelist$term_freqs)
```

### b. visualize with LDAvis
```{r}
library(LDAvis)
LDAvis::serVis(json)
```
